Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-01-31 analysis/paper_draft3.Rmd
restructure incompatibility piece in intro

Relying on historical strike data, however, requires an approach to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity). As a result, there may be incompatibility between plausible values of any attorney's bias parameter based on strike data for an attorney in the current trial and such values based on historical strike data for that attorney.

@@

We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges. We built this app with `r R.version$version.string` and the shiny package [@R; @shiny]. Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution. To show how lawyers might use this app in real cases in real time during selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017. Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later de-identified them. For details on this dataset, see the Appendix.

>>>>

Relying on historical strike data, however, requires an approach to  dealing with incompatibility between current and historical trail data. Such incompatibility might arise due to missingness in the historical data or to the possibility that attorney's strike bias depends on certain trial-level characteristics that vary across trials (e.g., defendant race, charge severity). As a result, incompatibility might cause an attorney's bias parameter based on strike data for an attorney in the current trial to differ from such values based on historical strike data for that attorney.

@@

We describe here a prototype software application ("app") that implements the approach described above and that attorneys and others can use in real time to detect bias in the use of peremptory challenges. We built this app with `r R.version$version.string` and the shiny package [@R; @shiny]. Unlike the simulations, we built the app using Rcpp [@Rcpp] and the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution. As in the simulations, a normal distribution was assumed for the distribution of $b$. The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in. No thinning was performed, as the correlation was weak and convergence occured rapidly. Convergence was checked from both traceplots and the Gelman-Rubin statistic [@Gelman1992]. Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (quicker with current data only _versus_ current plus historical data).

To show how lawyers might use this app in real cases in real time during selection, we loaded this app with actual strike data from a convenience sample of attorneys who appeared during jury selection in criminal cases in the federal district court for Connecticut during fiscal years 2013 through 2017. Using publicly-available court docket sheets for each case, we matched particular lawyers to these strikes and later de-identified them. For details on this dataset, see the Appendix.
